Visual Diagnostics of a Model Explainer -- Tools for the Assessment of LIME Explanations

### 2019-06-06 papers/Visual Diagnostics of a Model Explainer/visual-diagnostics-of-a-model-explainer.Rnw
worked more on paper to prepare for meeting

The European Union has taken the desire to provide explanations from machine learning models a step further. In May 2018, the General Data Protection Regulation (GDPR) went into effect (https://eugdpr.org/the-regulation/gdpr-faqs/). Part of the policies enforced by the regulation include the right for subjects affected by decisions made via automated algorithms to understand the logic behind the decisions being made. As Goodman and Flaxman (2016) point out, this regulation will have a great effect on the way that machine learning algorithms are used to make decisions. While there are different ways to interpret the ability to understand the logic behind the decisions, they claim a minimum would be to explain how the input variables relate to the predictions.

One approach to remedy the problem of uninterpretable models is to develop a model explainer that can provide insight into the performance of the complex model. A model explainer is a method that is separate from the model but uses the model and output from the model to provide understanding into mechanism of how the model produces predictions. (try to think of a simple example of a model explainer) (mention some of the model explainers that have been developed recently)

LIME (local interpretable model explanations) is one such model explainer (reference). While some model explainers are focused on understanding a model at the global level, LIME was designed to explain the model at the local level for a single prediction of interest. Conceptually, LIME fits a simple interpretable model, referred to as the explainer, that is meant to capture the behavior of the complex "black box model" in a local region around a prediction of interest. (a good place to include an image to help with this explanation)

The explanation that is produced from the explainer is also a prediction or an element of a model. Since it is not directly extracted from the complex model, it is important to ask the question, ``How reliable is the explanation?". Currently, there are no recommendations for how to assess the explanations from LIME (need to check to make sure this is a true statement). In this paper, we will present some tools that could help answer that question for LIME. Additionally, we will show an example using these tools on a forensics bullet matching dataset, and we will discuss how the results from the example provide some possible insights into the workings of LIME.

@@

\subsection{Application of LIME to Bullet Matching Data}

\subsubsection{lime R Package Perturbation Creation Methods}

@@

\subsubsection{Proposed Bin Creation Methods}

\section{Data}

>>>>

The European Union has taken the desire to provide explanations from machine learning models a step further. In May 2018, the General Data Protection Regulation (GDPR) went into effect (https://eugdpr.org/the-regulation/gdpr-faqs/). The regulation includes a policy that gives the right for subjects affected by decisions made via automated algorithms to understand the logic behind the decisions being made. As Goodman and Flaxman (2016) point out, this regulation will have a great effect on the way that machine learning algorithms are used to make decisions. At minimum, it will require that an explanation of how the input variables relate to the predictions.

One approach to remedy the problem of uninterpretable models is to develop a model explainer that can provide insight into the performance of the complex model. A model explainer is a method that is separate from the model but uses the model and output from the model to provide understanding into mechanism of how the model produces predictions. Recently, a handful of model explainers have been developed (reference). LIME (local interpretable model explanations) is one such model explainer (reference).

While some model explainers are focused on understanding a model at the global level, LIME was designed to explain the model at the local level for a single prediction of interest. Conceptually, LIME fits a simple interpretable model, referred to as the explainer, that is meant to capture the behavior of the complex "black box model" in a local region around a prediction of interest. (a good place to include an image to help with this explanation)

The explanation that is produced from the explainer is also a prediction or an element of a model. Since it is not directly extracted from the complex model, it is important to ask the question, ``How reliable is the explanation?". Currently, there are no recommendations for how to assess the explanations from LIME (need to check to make sure this is a true statement). In this paper, we will present some visualizations tools that can be used to assess the explanations from LIME in classification problems. Additionally, we will show an example using these tools on explanations from a forensics bullet matching dataset, and we will discuss how the results from the example provide some possible insights into the workings of LIME.

@@

\subsection{Implemention of LIME in R}

(section describing the procedure that is currently used by the LIME R package)

@@

\subsection{Visualization Diagnostic Tools for LIME}

\section{Application of LIME to Bullet Matching Data}
